---
title: "Reef Fishes - PELD ILOC"
output: 
  flexdashboard::flex_dashboard:
    social: menu
    navbar:
      - { icon: "fa-home", href: "https://peld-iloc.github.io/dashboards_peld-iloc/", align: right }
 
    
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(readr)
library(reshape2)
library(leaflet)
library(ggplot2)
library(vegan)
library(plotly)
library(lubridate)
library(dplyr)
library(RColorBrewer)
library(magrittr)
library(fontawesome)

```

```{r dataread, include = FALSE}

DF<- read.csv("../data/peld_iloc_peixes_recifais_Thiago_Silveira_2020_09_21.csv", header=T, sep=",", dec=".", na.strings = "NA", stringsAsFactors = T) 

glimpse(DF)
DF = as_tibble(DF)
DF$island<-DF$ï..island

DF = DF %>%
  select(-ï..island) %>% 
  relocate(island, .before = site)
  
DF


```

```{r taxonrank data, include = FALSE}
library(worms)

DF.taxa = DF %>% 
  mutate(species_name = str_replace_all(species_name, pattern = "_", replacement = " ")) %>% 
  mutate(species_name = str_to_sentence(species_name, locale = "pt")) %>% 
  mutate(verbatimIdentification = species_name) %>% 
  mutate(species_name = plyr::mapvalues(species_name,
                                        from = c("Platybelone argalus", "Labrisomus kalisherae", "Nicholsina usta usta", "Malacoctenus sp1", "Malacoctenus sp2", "Malacoctenus sp3", "Dasyatis americana", "Kyphosus incisor", "Emblemariopsis signifera", "Diplodus argenteus argenteus"),
                                        to = c("Platybelone argalus argalus", "Gobioclinus kalisherae", "Nicholsina usta", "Malacoctenus brunoi", "Malacoctenus lianae", "Malacoctenus lianae", "Hypanus americanus", "Kyphosus vaigiensis", "Emblemariopsis signifer", "Diplodus argenteus")))


sp_Names = unique(DF.taxa$species_name)
scNames_rank = wormsbynames(sp_Names, verbose = F)

# Kyphosus spp                                         no match                                 
# Menaphorus punticulatus                              no match                                 
# Entomacrodus spp                                     no match                                 
# Coryphopterus spb                                    no match                                 
# Sparisoma spp                                        no match                                 
# Ophioblennius spp                                    no match

# Changing names in DF.taxa
DF.taxa = DF.taxa %>%
    mutate(species_name = str_replace_all(species_name, pattern = "Coryphopterus spb", replacement = "Coryphopterus")) %>% # genus
    mutate(species_name = str_replace_all(species_name, pattern = "Entomacrodus spp", replacement = "Entomacrodus")) %>%  # genus
    mutate(species_name = str_replace_all(species_name, pattern = "Kyphosus spp", replacement = "Kyphosus")) %>% # genus
    mutate(species_name = str_replace_all(species_name, pattern = "Ophioblennius spp", replacement = "Ophioblennius")) %>% # genus
    mutate(species_name = str_replace_all(species_name, pattern = "Sparisoma spp", replacement = "Sparisoma")) # genus

# running again
sp_Names = unique(DF.taxa$species_name)
scNames_rank = wormsbynames(sp_Names, verbose = F)

# Not in worms database 
DF.taxa %>%
  filter(species_name == "Menaphorus punticulatus") # a hybrid between Paranthias furcifer e Cephalopholis fulva

# Create df for full join
name_cd_rk = scNames_rank %>% 
  select(scientificname, lsid, rank, AphiaID) %>% 
  filter(!is.na(rank))

DF.taxa = DF.taxa %>%
  full_join(name_cd_rk, by = c("species_name" = "scientificname"))

# Checking the existence of issues represented as "NA"
# As expected just for Menaphorus punticulatus, without lsid
DF.taxa %>%
  filter(is.na(lsid) == TRUE) %>% 
  distinct(species_name)

# Filling rank for unmatched species   
DF.taxa = DF.taxa %>%
  mutate_at(vars(rank), funs(ifelse(species_name == "Menaphorus punticulatus", "Species", . )))

```



General Information 
======================================================================

Column {data-width=350}
----------------------------------------------------------------------

### Zoom in to Explore PELD-ILOC Sites

```{r map}
## get lat lon
siteCoords = DF %>% dplyr::group_by(site,island)%>% 
  dplyr::summarise(lng = mean(lon_DD, na.rm=T), 
            lat = mean(lat_DD, na.rm=T))

## get total abund by site
siteAbund = DF %>%  
  filter(code != "NA" & abun!= "0")%>% 
  dplyr::group_by(site) %>% 
  dplyr::summarise(sumAbund = sum(abun, na.rm=T),
                   richness = n_distinct(code),
                   n_census = n_distinct(transect_id))


## add abund and to coords
siteCoords = full_join(siteCoords, siteAbund)


## make leaflet map
leaflet(siteCoords) %>% 
  setView(lng = -35, lat = -10, zoom = 04) %>% 
  addTiles() %>% 
  addCircleMarkers(label = ~paste0(str_to_upper(island),", ", site," / N. Taxa: ", richness, "/ N. Census: ",  n_census), 
                   radius = ~100 * (sumAbund/(sumAbund))/4,
                   fillOpacity = 0.9,
                   stroke = T,
                   clusterOptions = markerClusterOptions()) %>% 
  addMiniMap(toggleDisplay = T)

```




Row {data-width=350 data-padding=10}
----------------------------------------------------------------------

### Indicators

```{r value boxes, fig.align="center", echo = FALSE, fig.width=10, fig.height=7}
library(fontawesome)
library(emojifont)

  n_census = n_distinct(DF.taxa$transect_id)
  n_sites = n_distinct(DF.taxa$site)
  n_obs = n_distinct(DF.taxa$observer)
  n_sp = n_distinct(DF.taxa$species_name) 
  n_ind= sum(DF.taxa$abun)   
  max_depth = max(DF.taxa$depth_m)  



df <- data.frame(
    x = rep(seq(2, 15, 6.5), 2),
    y = c(rep(6.5, 3), rep(2,3)),
    h = rep(4.25, 6),
    w = rep(6.25, 6),
    value = as.vector(c(n_census, n_sites, n_obs, n_sp, n_ind, paste0(max_depth," ","m"))),
    
    info=c("Census", "Sites", "Observers", "Species", "Individuals", "Max depth"),
    
    
   shape = c(fontawesome("fa-clipboard"),
              fontawesome("fa-map-marker"),
              fontawesome("fa-user"),
              fontawesome("fa-barcode"),
              fontawesome("fa-list-ol"),
              fontawesome("fa-arrow-down")),
   
    
    font_family = c(rep("fontawesome-webfont", 6)),
    color = factor(1:6)
)


p2 <- ggplot(df, aes(x, y, height = h, width = w, label = info)) +
    geom_tile(aes(fill = color)) +
    geom_text(color = "white", fontface = "bold", size = 25,
              aes(label = value, x = x - 2.9, y = y + 1), hjust = 0) +
    geom_text(color = "white", fontface = "bold", size = 15,
              aes(label = info, x = x - 2.9, y = y - 1), hjust = 0) +
    coord_fixed() +
    scale_fill_brewer(type = "qual",palette = "Dark2") +
    geom_text(size = 50, aes(label = shape, family = font_family,
                             x = x + 1.5, y = y + 0.1), alpha = 0.25) +
    theme(plot.margin = margin(-1,-1,-1,-1, "cm"),
          panel.background = element_rect(fill =  "white"),
          line = element_blank(),
          text = element_blank(),
          title = element_blank()
          ) +
    #theme(plot.margin = margin(0,0,0,0, "cm")) +
    guides(fill = "none") 

p2


```

### Taxonomic rank distribution

```{r donut }


## summarizes the results
taxranks = as.data.frame(table(DF.taxa$rank))

## make a donut
p = taxranks %>% plot_ly(labels = ~Var1, values=~Freq, sizes = I(2)) %>% 
  add_pie(hole=0.6)
  # %>% 
  #layout(title = ~paste0("Total Taxa Number: ", length(sp_Names))) 

plotly::config(p,displayModeBar = F) 



```






Taxon List
======================================================================


Row 
-----------------------------------------------------------------------



### Taxa list - Number of transects each taxa where found. To access the taxonomic information about the species found, go to [Word Register of Marine Species](http://www.marinespecies.org/index.php), insert in the search field the corresponding AphiaID number to explore the complete taxa record. All taxa name were checked via Worms data base.  

```{r taxalist}

taxaTable = DF.taxa %>% 
  dplyr::group_by(species_name) %>%
  dplyr::summarise(
            AphiaID = unique(AphiaID),
            Taxon = unique(species_name),  
            "Noronha" = sum(island =="noronha", na.rm=T), 
            "Atol das Rocas" = sum(island=="rocas", na.rm=T), 
            "ASPS" = sum(island=="stpauls_rocks", na.rm=T), 
            "Trindade" = sum(island=="trindade", na.rm=T)) %>%
  select("AphiaID", "Taxon",	"Noronha",	"Atol das Rocas",	"ASPS",	"Trindade")


taxaTable = taxaTable %>% 
  mutate(Total = apply(.[3:6], 1, sum)) %>% 
  arrange(desc(Total))

knitr::kable(taxaTable)

```


Collector Curves {.storyboard}
=======================================================

### **Fernando de Noronha** - Species accumulation curves by site 


```{r rarenoronha}
## filter by island
Occurrence.site = DF.taxa %>% filter(island=="noronha")

# do it by site
siteNames = unique(Occurrence.site$site)

## empty DF for results
sppAccum = data.frame(site = character(),
                      ntrans = numeric(),
                      richness = numeric(), 
                      sd = numeric())

for (i in 1:length(siteNames)){
  ecoMat = dcast(transect_id~species_name, data=subset(Occurrence.site, site==siteNames[i]), 
                 value.var = "AphiaID", length)
  
  sppcurve = specaccum(ecoMat[,-1], method = "random")
  sppAccum = rbind(sppAccum, 
                   data.frame(site = rep(siteNames[i], length(sppcurve$sites)),
                              ntrans = sppcurve$sites, 
                              richness = sppcurve$richness,
                              sd = sppcurve$sd))
}



## make the plot
pp = ggplot(sppAccum, aes(ntrans, richness, fill = site, colour = site))
pp = pp + geom_ribbon(aes(ymin=richness-sd, ymax=richness+sd), alpha = 0.3) + geom_line() + geom_point() + 
  theme_bw(base_size = 10) + xlab("Number of transects") + ylab("Accumulated number of Species")

ggplotly(pp) %>% plotly::config(displayModeBar = F) 

```

### **Atol das Rocas** - Species accumulation curves by site 


```{r rarerocas}
## filter by island
Occurrence.site = DF.taxa %>% filter(island=="rocas")

# do it by site
siteNames = unique(Occurrence.site$site)

## empty DF for results
sppAccum = data.frame(site = character(),
                      ntrans = numeric(),
                      richness = numeric(), 
                      sd = numeric())

for (i in 1:length(siteNames)){
  ecoMat = dcast(transect_id~species_name, data=subset(Occurrence.site, site==siteNames[i]), 
                 value.var = "AphiaID", length)
  
  sppcurve = specaccum(ecoMat[,-1], method = "random")
  sppAccum = rbind(sppAccum, 
                   data.frame(site = rep(siteNames[i], length(sppcurve$sites)),
                              ntrans = sppcurve$sites, 
                              richness = sppcurve$richness,
                              sd = sppcurve$sd))
}

##Add custon pallete, same color showed in map
##change this analysis using INext


## make the plot
pp = ggplot(sppAccum, aes(ntrans, richness, colour = site, fill = site))
pp = pp + geom_ribbon(aes(ymin=richness-sd, ymax=richness+sd), alpha=0.3) + geom_line() + geom_point() + 
  theme_bw(base_size = 10) + xlab("Number of transects") + ylab("Accumulated number of Species")

ggplotly(pp) %>% plotly::config(displayModeBar = F) 

```

### **ASPS** - Species accumulation curves by site 


```{r rareASPS}
## filter by island
Occurrence.site = DF.taxa %>% filter(island=="stpauls_rocks")

# do it by site
siteNames = unique(Occurrence.site$site)

## empty DF for results
sppAccum = data.frame(site = character(),
                      ntrans = numeric(),
                      richness = numeric(), 
                      sd = numeric())

for (i in 1:length(siteNames)){
  ecoMat = dcast(transect_id~species_name, data=subset(Occurrence.site, site==siteNames[i]), 
                 value.var = "AphiaID", length)
  
  sppcurve = specaccum(ecoMat[,-1], method = "random")
  sppAccum = rbind(sppAccum, 
                   data.frame(site = rep(siteNames[i], length(sppcurve$sites)),
                              ntrans = sppcurve$sites, 
                              richness = sppcurve$richness,
                              sd = sppcurve$sd))
}

##Add custon pallete, same color showed in map
##change this analysis using INext


## make the plot
pp = ggplot(sppAccum, aes(ntrans, richness, colour = site, fill = site))
pp = pp + geom_ribbon(aes(ymin=richness-sd, ymax=richness+sd), alpha=0.3) + geom_line() + geom_point() + 
  theme_bw(base_size = 10) + xlab("Number of transects") + ylab("Accumulated number of Species")

ggplotly(pp) %>% plotly::config(displayModeBar = F) 

```

### **Trindade** - Species accumulation curves by site 


```{r raretrindade}
## filter by island
Occurrence.site = DF.taxa %>% filter(island=="trindade")

# do it by site
siteNames = unique(Occurrence.site$site)

## empty DF for results
sppAccum = data.frame(site = character(),
                      ntrans = numeric(),
                      richness = numeric(), 
                      sd = numeric())

for (i in 1:length(siteNames)){
  ecoMat = dcast(transect_id~species_name, data=subset(Occurrence.site, site==siteNames[i]), 
                 value.var = "AphiaID", length)
  
  sppcurve = specaccum(ecoMat[,-1], method = "random")
  sppAccum = rbind(sppAccum, 
                   data.frame(site = rep(siteNames[i], length(sppcurve$sites)),
                              ntrans = sppcurve$sites, 
                              richness = sppcurve$richness,
                              sd = sppcurve$sd))
}

##Add custon pallete, same color showed in map
##change this analysis using INext


## make the plot
pp = ggplot(sppAccum, aes(ntrans, richness, colour=site, fill = site))
pp = pp + geom_ribbon(aes(ymin=richness-sd, ymax=richness+sd), alpha=0.3) + geom_line() + geom_point() + 
  theme_bw(base_size = 10) + xlab("Number of transects") + ylab("Accumulated number of Species")

ggplotly(pp) %>% plotly::config(displayModeBar = F) 

```




Taxa abundance {.storyboard}
=======================================================

### **Fernando de Noronha** - Number of individuals recorded by site (individuals/ n. census). The plots shows the Top 10 most abundant taxa by site.

```{r abun noronha}

taxafreq = DF.taxa %>% 
  dplyr::filter( island == "noronha" & !is.na(AphiaID) & abun!="0" & species_name!="NA") %>% #ok
  dplyr::mutate(site = str_replace_all(site, pattern = "_", replacement = " ")) %>% 
  dplyr::mutate(site_code = abbreviate(site)) %>%             
  dplyr::group_by(island, site, site_code, species_name) %>%
  
  dplyr::summarise(sppfreq = n(),
                   sumAbund = sum(abun, na.rm=T),
                   n_census = n_distinct(transect_id)) %>% 
  dplyr::mutate(abun_std = sumAbund/n_census) %>% 
  dplyr::arrange(desc(abun_std)) %>% 
  dplyr::mutate(sppacum = cumsum(sppfreq)) %>%  
  dplyr::top_n(10, sppfreq) # top ten taxa



pp = ggplot(taxafreq, aes(species_name, abun_std, fill = site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site_code) + 
  theme_bw(base_size = 9) + xlab("") + ylab("ind/ n. visual census") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 
```

### **Atol das rocas** - Number of individuals recorded by site (individuals/ n. census). The plots shows the Top 10 most abundant taxa by site.

```{r abun rocas}

taxafreq = DF.taxa %>% 
  dplyr::filter( island == "rocas" & !is.na(AphiaID) & abun!="0" & species_name!="NA") %>% #ok
  dplyr::mutate(site = str_replace_all(site, pattern = "_", replacement = " ")) %>% 
  dplyr::mutate(site_code = abbreviate(site)) %>%             
  dplyr::group_by(island, site, site_code, species_name) %>%
  
  dplyr::summarise(sppfreq = n(),
                   sumAbund = sum(abun, na.rm=T),
                   n_census = n_distinct(transect_id)) %>% 
  dplyr::mutate(abun_std = sumAbund/n_census) %>% 
  dplyr::arrange(desc(abun_std)) %>% 
  dplyr::mutate(sppacum = cumsum(sppfreq)) %>%  
  dplyr::top_n(10, sppfreq) # top ten taxa



pp = ggplot(taxafreq, aes(species_name, abun_std, fill = site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site_code) + 
  theme_bw(base_size = 9) + xlab("") + ylab("ind/ n. visual census") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 
```

### **ASPSP** - Number of individuals recorded by site (individuals/ n. census). The plots shows the Top 10 most abundant taxa by site.

```{r abun ASPSP}

taxafreq = DF.taxa %>% 
  dplyr::filter( island == "stpauls_rocks" & !is.na(AphiaID) & abun!="0" & species_name!="NA") %>% #ok
  dplyr::mutate(site = str_replace_all(site, pattern = "_", replacement = " ")) %>% 
  dplyr::mutate(site_code = abbreviate(site)) %>%             
  dplyr::group_by(island, site, site_code, species_name) %>%
  
  dplyr::summarise(sppfreq = n(),
                   sumAbund = sum(abun, na.rm=T),
                   n_census = n_distinct(transect_id)) %>% 
  dplyr::mutate(abun_std = sumAbund/n_census) %>% 
  dplyr::arrange(desc(abun_std)) %>% 
  dplyr::mutate(sppacum = cumsum(sppfreq)) %>%  
  dplyr::top_n(10, sppfreq) # top ten taxa



pp = ggplot(taxafreq, aes(species_name, abun_std, fill = site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site_code) + 
  theme_bw(base_size = 9) + xlab("") + ylab("ind/ n. visual census") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 

```
### **Trindade** - Number of individuals recorded by site (individuals/ n. census). The plots shows the Top 10 most abundant by site.

```{r abun trindade}

taxafreq = DF.taxa %>% 
  dplyr::filter( island == "trindade" & !is.na(AphiaID) & abun!="0" & species_name!="NA") %>% #ok
 dplyr::mutate(site = str_replace_all(site, pattern = "_", replacement = " ")) %>% 
  dplyr::mutate(site_code = abbreviate(site)) %>%             
  dplyr::group_by(island, site, site_code, species_name) %>%
  
  dplyr::summarise(sppfreq = n(),
                   sumAbund = sum(abun, na.rm=T),
                   n_census = n_distinct(transect_id)) %>% 
  dplyr::mutate(abun_std = sumAbund/n_census) %>% 
  dplyr::arrange(desc(abun_std)) %>% 
  dplyr::mutate(sppacum = cumsum(sppfreq)) %>%  
  dplyr::top_n(10, sppfreq) # top ten taxa



pp = ggplot(taxafreq, aes(species_name, abun_std, fill = site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site_code) + 
  theme_bw(base_size = 9) + xlab("") + ylab("ind/ n. visual census") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 

```