---
title: "Reef Fishes - PELD ILOC"
output: 
  flexdashboard::flex_dashboard:
    social: menu
    navbar:
      - { icon: "fa-home", href: "https://peld-iloc.github.io/dashboards_peld-iloc/", align: right }
 
    
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(readr)
library(reshape2)
library(leaflet)
library(ggplot2)
library(vegan)
library(plotly)
library(lubridate)
library(dplyr)
library(RColorBrewer)
library(magrittr)
library(fontawesome)

```

```{r dataread, include = FALSE}

rm(list = ls())
baseDir = "C:/Users/silve/OneDrive/Documentos/Academico/POS-DOC_PELDILOC/dashboards_peld-iloc/docs/data"

#"C:/Users/silve/OneDrive/Documentos/Academico/POS-DOC_PELDILOC/dashboards_peld-iloc/docs/data"


file_sites = "bentos_todas.csv"
file_coords = "Peld_Site_Coordinates_2021_07_12.csv"

dados <- read.csv(file.path(baseDir, file_sites), header = TRUE) %>% 
  select(-X) %>% 
  mutate(categoryid = recode("POR " = "POR", categoryid))


DF.coords = read_delim(file.path(baseDir, file_coords),
                       na="NA",
                       delim = ",",
                       skip_empty_rows = TRUE)

# check unique ids
dados %>% 
  distinct(categoryid) %>% 
  arrange(categoryid)


```

```{r taxa, include=FALSE}
dados <- dados %>% 
  mutate(categoryname = plyr::mapvalues(categoryname, 
                                        from = c("jania", "amphiroa_sp_jania_sp", "outra_macroalga_calcaria_articulada", "amphiroa_sp_jania_sp2N", "outra_macroalga_calcaria_articulada_2N", "outra_macroalga_calcaria_articulada_1N", "amphiroa_sp_jania_spN", "outra_macroalga_calcaria_articulada_3N", 
                                                 "turf_calcareo", "encrusting_crustose_algae", "outra_macroalga_crostosa_3", "outra_macroalga_crostosa_4", "rhodophyta", "outra_macroalga_crostosa_2","outra_macroalga_crostosa_1", "calcarias_nao_articuladas_calcaria_crostosa", "outra_macroalga_crostosa_1R", "outra_macroalga_crostosa_1N", "outra_macroalga_crostosa_2N", "outra_macroalga_crostosa_3N", "outra_macroalga_crostosa_4N",
                                                 "peyssonelia", "estrelas", 
                                                 "cianobacteria", "cianobacterias_1R", "cianobacterias_3R", "cianobacterias_2R", "cianobacteria_1N", "cianobacteria_2N", "cianobacteria_3N", "cianobacteria_4N",
                                                 "chlorophyta", "outra_macroalga_filamentosa_1R", "outra_macroalga_filamentosa_2R", "outra_macroalga_filamentosa_3R", "outra_macroalga_filamentosa_4R", "outra_macroalga_filamentosa_1N", "outra_macroalga_filamentosa_2N", "outra_macroalga_filamentosa_3N",
                                                 "bryopsis_pennata", 
                                                 "dictyota_sp1R", "dictyota_ciliolata", "dictyota_sp2R", "dictyota_menstrualis", "dictyota_sp3R", "dictyota_spp_1N", "dictyota_spp_4N", "dictyota_spp_2N", "dictyota_spp_7N", "dictyota_spp_6N", "dictyota_spp_3N", "canistrocarpus_cervicornis", "dictyota_spp_5N", "dictyota_pulchella",
                                                 "outra_macroalga_foliaceaR", "outra_macroalga_foliacea_1N",
                                                 "padina_gymnospora",
                                                 "dictyopteris_justii", "dictyopteris_spp_1N", "dictyopteris_spp_3N", "dictyopteris_spp_2N", "dictyopteris_jolyana",
                                                 "incrustrante_2t", "incrustante_1R", "incrustante_2R", "incrustante_3R", "incrustrante_2N", "incrustrante_3N", "outra_esponja_2N", "outra_esponja_4N", "outra_esponja_1N", "placospongia", "incrustrante_1N", "outra_esponja_3N", "tubular",
                                                 "coral_1", 
                                                 "milepora_incrustante", "milepora",
                                                 "caulerpa_verticilata",
                                                 "stoloniferous_algae",
                                                 "zoanthus_sociatus",
                                                 "palythoa_caribeorum",
                                                 "siderastrea_stellata",
                                                 "red_filamentous_turf", "corticated_turf", "turf_filamentoso", "outro_turf"),
                                        to = c(rep("articulated_coralline_algae", 8),
                                               rep("CCA", 13),
                                               "peyssonnelia", "Asteroidea",
                                               rep("cyanobacteria", 8),
                                               rep("filamentous_algae", 8),
                                               "bryopsis", 
                                               rep("dictyota", 14),
                                               rep("foliose_algae", 2), 
                                               "padina",
                                               rep("dictyopteris", 5),
                                               rep("porifera", 13), 
                                               "scleractinia",
                                               rep("millepora", 2),
                                               "caulerpa_verticillata",
                                               "caulerpa_racemosa_ASPSP",
                                               "zoanthid",
                                               "palythoa",
                                               "siderastrea",
                                               rep("turf", 4))),
         categoryid = ifelse(categoryid == "SAR", "COR", categoryid)) 

#####
dados1 <- dados %>% 
  mutate(unique_id = paste0(island, sites, year, depth, transect, quadrate, image)) %>% 
  mutate(unique_trans_id = paste0(island, sites, year, depth, transect)) %>% 
  select(island, sites, depth_mean, year, unique_trans_id, unique_id, categoryname, categoryid, coverpercategory) %>% 
  mutate(unique_id = recode(unique_id, "stpauls_rocksenseada2015deep832014_T8_3.jpg" = "stpauls_rocksenseada2014deep832014_T8_3.jpg",
                            "stpauls_rocksenseada2016deep832014_T8_3.jpg" = "stpauls_rocksenseada2014deep832014_T8_3.jpg",
                            "stpauls_rocksenseada2016deep882015_T8_8.JPG" = "stpauls_rocksenseada2015deep882015_T8_8.JPG")) %>% 
  group_by(island, sites, depth_mean, year, unique_trans_id, unique_id, categoryname, categoryid) %>% 
  dplyr::summarise(coverpercategory = sum(coverpercategory)) %>% 
  filter(unique_id != "trindadePonta Norte2013Fundo3Quadrat6trindade_Ponta Norte_2013_10_Quadrat6") %>% 
  mutate(categoryname = recode(categoryname, "CCA" = "Lithophylloideae",
                               "articulated_coralline_algae" = "Amphiroideae",
                               "desconhecido" = "unknown",
                               "areia_cascalho" = "sand_gravel",
                               "bivalve_1" = "bivalvia",
                               "zoanthid" = "Zoantharia",
                               "outra_macroalga_corticadaR" = "corticated",
                               "outra_macroalga_coriaceaR" = "leathery",
                               "foliose_algae" = "foliose",
                               "filamentous_algae" = "filamentous",
                               "green_filamentous_algae" = "filamentous",
                               "caulerpa_racemosa_ASPSP" = "caulerpa_racemosa"))

```


```{r data coords, include=FALSE}

# Joining the tables using as reference the site (DF.sites) and site_coords (DF.coords)
DF.benthos <- dados1 %>%
  mutate(sites = str_to_lower(sites, locale = "pt")) %>%
  mutate(sites = recode(sites, "farrilhoes" = "farilhoes", "ponta norte" = "ponta_norte")) %>%
  left_join(DF.coords, by = c("sites" = "site_coord"), copy = TRUE) %>% 
  mutate(island = paste(toupper(substr(island, 1, 1)), substr(island, 2, nchar(island)), sep = ""))


# Reorganizing DF.sites with the corrected coordinates and site names
DF.benthos1 <- DF.benthos %>% 
  data.frame() %>% 
  dplyr::select(-sites, -checking_status, -Obs) %>% 
  relocate(New_name_coord, .after = island) %>% 
  dplyr::rename(site = New_name_coord,
                lon_DD = lon_DD_coord,
                lat_DD = lat_DD_coord) %>% 
  mutate(categoryname = recode(categoryname, "porites_asteroides" = "porites_astreoides",
                               "xetospongia_muta" = "xestospongia_muta",
                               "hypnea_musciformes" = "hypnea_musciformis",
                               "diadema_antillarium" = "diadema_antillarum"))

```






General Information 
======================================================================

Column {data-width=350}
----------------------------------------------------------------------

### Zoom in to Explore PELD-ILOC Sites

```{r map}


## get lat lon
siteCoords = DF.benthos1 %>% dplyr::group_by(site,island)%>% 
  dplyr::summarise(lng = mean(lon_DD, na.rm=T), 
            lat = mean(lat_DD, na.rm=T))

## get total abund by site
site_rich = DF.benthos1 %>%  
  filter(categoryid != "NA" & coverpercategory!= 0)%>% 
  dplyr::group_by(site) %>% 
  dplyr::summarise(richness = n_distinct(categoryid),
                   n_quadrats = n_distinct(unique_id)
                   )


## add abund and to coords
siteCoords = full_join(siteCoords, site_rich)


## make leaflet map
leaflet(siteCoords) %>% 
  setView(lng = -35, lat = -5, zoom = 06) %>% 
  addTiles() %>% 
  addCircleMarkers(label = ~paste0(str_to_upper(island),", ", site," / N. categories: ", richness, "/ N. quadrats: ",  n_quadrats), 
                   radius = ~100 * (richness/80),
                   fillOpacity = 0.9,
                   stroke = T,
                   clusterOptions = markerClusterOptions()) %>% 
  addMiniMap(toggleDisplay = T)

```




Row {data-width=350 data-padding=10}
----------------------------------------------------------------------

### Indicators

```{r value boxes, fig.align="center", echo = FALSE, fig.width=10, fig.height=7}

library(fontawesome)
library(emojifont)

  n_transects = n_distinct(DF.benthos1$unique_trans_id)
  n_sites = n_distinct(DF.benthos1$site)
  n_cat = n_distinct(DF.benthos1$categoryname) 
  
  

  
  

df <- data.frame(
    x = seq(2, 15, 6.5),
    y = rep(6.5, 3),
    h = rep(4.25, 3),
    w = rep(6.25, 3),
    value = as.vector(c(n_transects, n_sites, n_cat)),
    info=c("Transects", "Sites", "Categories"),
    shape = c(fontawesome("fa-clipboard"),
              fontawesome("fa-map-marker"),
              fontawesome("fa-barcode")),
    
    font_family = c(rep("fontawesome-webfont", 3)),
    color = factor(1:3)
              )


p2 <- ggplot(df, aes(x, y, height = h, width = w, label = info)) +
    geom_tile(aes(fill = color)) +
    geom_text(color = "white", fontface = "bold", size = 25,
              aes(label = value, x = x - 2.9, y = y + 1), hjust = 0) +
    geom_text(color = "white", fontface = "bold", size = 15,
              aes(label = info, x = x - 2.9, y = y - 1), hjust = 0) +
    coord_fixed() +
    scale_fill_brewer(type = "qual",palette = "Dark2") +
    geom_text(size = 50, aes(label = shape, family = font_family,
                             x = x + 1.5, y = y + 0.1), alpha = 0.25) +
    theme(plot.margin = margin(-1,-1,-1,-1, "cm"),
          panel.background = element_rect(fill =  "white"),
          line = element_blank(),
          text = element_blank(),
          title = element_blank()
          ) +
    #theme(plot.margin = margin(0,0,0,0, "cm")) +
    guides(fill = "none") 

p2


```

### Categories distribution

```{r donut }


## summarizes the results
catranks = as.data.frame(table(DF.benthos1$categoryid))

## make a donut
p = catranks %>% 
  top_n(10) %>% 
  plot_ly(labels = ~Var1, values=~Freq, sizes = I(2)) %>% 
  add_pie(hole=0.6)
  # %>% 
  #layout(title = ~paste0("Total Taxa Number: ", length(sp_Names))) 

plotly::config(p,displayModeBar = F) 



```






Taxon List
======================================================================


Row 
-----------------------------------------------------------------------



### Category list - Number of transects each category where found.  

```{r taxalist}

taxaTable = DF.benthos1%>% 
  dplyr::group_by(categoryid, categoryname) %>%
  dplyr::summarise(
            "Category name" = unique(categoryname),
            "Category ID" = unique(categoryid),
            "Noronha" = sum(island =="Noronha", na.rm=T), 
            "Atol das Rocas" = sum(island=="Rocas", na.rm=T), 
            "ASPS" = sum(island=="Stpauls_rocks", na.rm=T), 
            "Trindade" = sum(island=="Trindade", na.rm=T)) %>%
  select("categoryname", "categoryid", 	"Noronha",	"Atol das Rocas",	"ASPS",	"Trindade")


taxaTable = taxaTable %>% 
  mutate(categoryname = ifelse(str_detect(categoryname, "^[:upper:]+$"), categoryname,str_to_title(categoryname))) %>% 
  mutate(categoryname = str_replace_all(categoryname, "_", " "))%>% 
  rowwise() %>% 
  mutate(Total = sum(across(Noronha:Trindade))) %>% 
  arrange(desc(Total))

knitr::kable(taxaTable)


```





Taxa abundance {.storyboard}
=======================================================

### **Fernando de Noronha** - Relative  cover recorded by site. 
```{r abun noronha}
#### solution
taxacover = DF.benthos1 %>% 
dplyr::select(island, site, categoryid, coverpercategory) %>% 
  dplyr::filter(island == "Noronha" & coverpercategory!= 0 & categoryid!="NA") %>% #ok
  dplyr::group_by(site, categoryid)%>% 
  dplyr::summarize(mean_mean = mean(coverpercategory),
                   sum_cover = sum(coverpercategory)) %>% 
  dplyr::mutate(sum_cover_site = sum(sum_cover)) %>% 
  dplyr::mutate(cover_site = (sum_cover*100)/sum_cover_site) %>% 
  dplyr::arrange(categoryid)


pp = ggplot(taxacover, aes(categoryid, cover_site, fill= site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site) + 
  theme_bw(base_size = 9) + xlab("") + ylab("relative cover") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 



```

### **Atol das rocas** - Relative  cover recorded by site. 

```{r abun rocas}
unique(DF.benthos1$island)

taxacover = DF.benthos1 %>% 
dplyr::select(island, site, categoryid, coverpercategory) %>% 
  dplyr::filter(island == "Rocas" & coverpercategory!= 0 & categoryid!="NA") %>% #ok
  dplyr::group_by(site, categoryid)%>% 
  dplyr::summarize(mean_mean = mean(coverpercategory),
                   sum_cover = sum(coverpercategory)) %>% 
  dplyr::mutate(sum_cover_site = sum(sum_cover)) %>% 
  dplyr::mutate(cover_site = (sum_cover*100)/sum_cover_site) %>% 
  dplyr::arrange(categoryid)


pp = ggplot(taxacover, aes(categoryid, cover_site, fill= site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site) + 
  theme_bw(base_size = 9) + xlab("") + ylab("relative cover") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 

```

### **ASPSP** - Relative  cover recorded by site. 

```{r abun ASPSP}

unique(DF.benthos1$island)

taxacover = DF.benthos1 %>% 
dplyr::select(island, site, categoryid, coverpercategory) %>% 
  dplyr::filter(island == "Stpauls_rocks" & coverpercategory!= 0 & categoryid!="NA") %>% #ok
  dplyr::group_by(site, categoryid)%>% 
  dplyr::summarize(mean_mean = mean(coverpercategory),
                   sum_cover = sum(coverpercategory)) %>% 
  dplyr::mutate(sum_cover_site = sum(sum_cover)) %>% 
  dplyr::mutate(cover_site = (sum_cover*100)/sum_cover_site) %>% 
  dplyr::arrange(categoryid)


pp = ggplot(taxacover, aes(categoryid, cover_site, fill= site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site) + 
  theme_bw(base_size = 9) + xlab("") + ylab("relative cover") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 


```
### **Trindade** - Relative  cover recorded by site. 

```{r abun trindade}

unique(DF.benthos1$island)

taxacover = DF.benthos1 %>% 
dplyr::select(island, site, categoryid, coverpercategory) %>% 
  dplyr::filter(island == "Trindade" & coverpercategory!= 0 & categoryid!="NA") %>% #ok
  dplyr::group_by(site, categoryid)%>% 
  dplyr::summarize(mean_mean = mean(coverpercategory),
                   sum_cover = sum(coverpercategory)) %>% 
  dplyr::mutate(sum_cover_site = sum(sum_cover)) %>% 
  dplyr::mutate(cover_site = (sum_cover*100)/sum_cover_site) %>% 
  dplyr::arrange(categoryid)


pp = ggplot(taxacover, aes(categoryid, cover_site, fill= site))
pp = pp + geom_bar(stat="identity") + coord_flip() + facet_grid(~site) + 
  theme_bw(base_size = 9) + xlab("") + ylab("relative cover") # ok no NA'S

ggplotly(pp) %>% plotly::config(displayModeBar = F) 


```